---
title: "Data Cleaning"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages used 
```{r, messages = FALSE}
library(rvest)
library(readr)
library(dplyr)
library(janitor)
library(ggplot2)
```

## Loading the data
```{r}
webpage <- read_html("https://www.pwrc.usgs.gov/bbl/manual/speclist.cfm")

tbls <- html_nodes(webpage, "table") %>% 
    html_table(fill = TRUE)

species <- tbls[[1]] %>% 
    clean_names() %>% 
    select(alpha_code, common_name) %>% 
    mutate(alpha_code = tolower(alpha_code))
```

janitor package helps clean up column names 
clean_names() - no special characters, lower case, no spaces 

```{r, message = F}
pred_original <- read_csv("https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3A9ffec04c-7e2d-41dd-9e88-b6c2e8c4375e")
nest_original <- read_csv("https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3A982bd2fc-4edf-4da7-96ef-0d11b853102d")
```

### Writing a function that will translate species code into common names
I want the function to grab the species common name from the species db and relate it to the alpha code 
```{r}
alpha_to_common <- function(dataframe, definitions) {
  return_dataframe <- left_join(dataframe, definitions, by = c("species" = "alpha_code"))
  return(return_dataframe)
}

pred <- alpha_to_common(dataframe = pred_original, definitions = species)
nest <- alpha_to_common(dataframe = nest_original, definitions = species)
```
## Alternative 

```{r}
#' Function to add common name to data.frame according to the BBL list of species codes

#' @param df A data frame containing BBL species codes in column `species`
#' @param species A data frame defining BBL species codes with columns `alpha_code` and `common_name`
#' @return A data frame with original data df, plus the common name of species


#assign_species_name <- function(df, species){
    if (!("alpha_code" %in% names(species)) |
        !("species" %in% names(df)) |
        !("common_name" %in% names(species))){
      stop("Tables appear to be formatted incorrectly.")
    }  
  
    return_df <- left_join(df, species, by = c("species" = "alpha_code"))
    
    if (nrow(return_df) > nrow(df)){
      warning("Joined table has more rows than original table. Check species table for duplicated code values.")
    }
    
    if (length(which(is.na(return_df$common_name))) > 0){
      x <- length(which(is.na(return_df$common_name)))
      warning(paste("Common name has", x, "rows containing NA"))
    }
    
    return(return_df)
        
}

```

### Total number oredators by year and species
```{r}
pred_year_sp <- pred %>%
  group_by(year, common_name) %>%
  summarize(Count = sum(count))%>%
  filter(!is.na(common_name))

ggplot(pred_year_sp, aes (x=common_name, y=Count, fill = common_name)) +
  geom_col()+
  facet_wrap(~year)
```
### Total number of eggs predated by year and species

```{r}
eggs_year_species <- nest %>%
  group_by(year, common_name) %>%
  summarise("Eggs_predated" = sum(number_eggs_predated))%>%
filter(!is.na(common_name))
```
### Total number of predators by year, join to summarise egg predation data, and plot 

```{r}
eggs_pred_year_species <- left_join(eggs_year_species, pred_year_sp, by = "year")

ggplot(eggs_pred_year_species, aes (x=Count, y=Eggs_predated, fill = common_name)) +
  geom_point()+
  facet_wrap(~common_name)
```

